Title here

### 2020-05-31 paper/paper.Rmd
add noisy index function section; add polish alpha for search_goedesic section

```{r load-data}
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
```

@@

Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data. A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine. The projection pursuit index measures the "interestingness" of data defined as its departure from normality. Numerous indices have been proposed in the literature, including lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection]. An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index. The discussion of existing optimisation procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path. An illustration modified from [@buja2005computational] entails the geodesic interpolation bewteen the plane generated by projection pursuit.

@@

As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the com- putational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe. An ideal optimisation procedure needs to have following characteristics:

- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentible

- *Being able to optimise with constraints*: the projection matrix is restricted to an orthornormal matrix.

@@

@posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis. The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$. The new basis is taken as the target basis if it has higher index value, or the sampling continues. If no basis is found to have higher index value after a certain humber of tries $n$, the radius $c$ is halved. The algorithm stops when the maximum numbere of iteration is attained or the radius $c$ is less than a pre-determined number. [Pursuit package uses this method and it works greate! But I don't think we implement this - although don't think it is too hard to do it].

@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable. Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood. Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.

@@

*Think about how does your pacakge help people to understand optimisation*

 - diagnostic on stochastic optim
 - vis the prograssion of multi-parameter decision variable 
 - understanding learning rate - neightbourhood parameter
 - understand where the local & global maximum is found - trace plot - see if noisy function

@@

The psudo-code below illustrates the implementation of guided tour in the tourr package. Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path. This information is then used to compute a series of interpolating bases inside the `tour()` function. All the basis will be sent to create animation for visualising the tour in the `animate()` function.

@@

Visualisatioin has been widely used for exploring and understanding data. Visualisation presents information in a graphicial manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation. Visual diagnostics can be real-time or post-run. Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.   This section focuses on the definition and production of post-run diagnoistics and the next section discusses real-time diagnostics.

@@

The graphical system in R is established based on the conception gramma of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.  A layer includes 1) the dataset that powers the plot; 2) a geometric object that represents the visual shape of the plot and 3) relevant statistical transformation if needed. An important concept in the gramma of graphic is *aesthetic mapping*. Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object. For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot. To create a boxplot, we first map one variable on the x-axis and then map the five point summary of another variable on the y-axis. This computation of five point summary from the origin variable is statistical transformation. This definition of graphic through layers provides advantages to produce complex plots since all the plots, however complex, can be decomposed into basic geometric objects shaped by the variable supplied.

@@

The concept of gramma of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm. The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest. The next section shows how a global object is created in projection pursuit guided tour.

@@

In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randonly nearby bases generated in the optimisation are not stored. This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[needa rephrase this part].**

Two set of simulated data are used in the demonstration of the visualiation and diagnostics of the tour optimisation. A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`). `x2` is the informative variable simulated from two bi-modal normal distribution centered at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`. The data has been scaled to ensure `x2` has variance of 1.

@@

Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure. The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process. The example below presents the global object of a 1D projection of the small dataset with geodesic seraching method.

@@

Given all the information of interest in a global object, the next thing is to design its visual representation that can diagnose our algorithm. In an iterative algorithm, one would be interested to know how one parameter changes as the algorithm iterates. The parameter could be could be the decision variable, the value of the optimisation function or other parameter that facilitate the algorithm. In projectio pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`). These are the variables that will be mapped into the diagnostic plot. The global object also has its time series nature since the observations are recorded as the algorithm progresses and thus can't be switch in rows. This time series structure is always hierachical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm. Take our example, in the global object, each row has an `id` label which is the smallest unit the observation is ordered. A larger ordering unit is `tries` since it is updated everytime a new target basis is found. A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`. These features explain the definition of the global object for projection pursuit guided tour in Figure \ref{fig:glb-obj}.

@@

```{r points-id, fig.asp = 0.5, fig.cap = "\\label{points-id}this is a xxx"}
dt <- holes_2d_better_max_tries %>%
  filter(info != "interpolation") %>% 
  mutate(id = row_number())

dt %>% 
  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
  geom_point()  +
  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
            aes(group = 1)) 
```

@@

@@

Plotting the interpolating points as time series data allows us to diagnose characterstics of different configurations and index functions. Here we present two examples of using  plots to diagnose the tour algorithm and different index functions.

@@

*Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index. As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to ahve a higher index value. After the interpolation, the target basis will become the current basis and send back to the projection pursuit algorithm to find the next target basis. From figure \ref{fig:interruption}, it is possible that there are bases with index value higher than the target basis on the interpolation path and these bases could be used to search for new basis in the next iteration.

Thus an interruption is constructed to accpet the interpolating bases up to the one with the larget index value on the interpolation path, and that basis is taken as the current basis for the next iteration. After implementing this interruption, the tracing plot with the same configuration is shown on the lower panel. We can observe that rather than interpolating to the target basis at `id = 62`, the interpolation stops before the index value starts to decrease at `id = 60`. This implementation results in a higher index value in the end with fewer steps.

@@

*Example: Noisy index function* Add an example on noisy function

<!-- The tracing plot is useful to explore the index value. Given a global object created, `explore_trace_all()` plots the index value of all the bases.  -->

xxx

Currently, I have three things to put for the noisy index function.

1) interpolating plot finds that search_geodesic is not doing a good job on optimising the noisy index; while search_better is.

2) The interpolating plot for noisy index function is not smooth but the target basis is always increasing.

3) search_better_random sees to find both local and global maximum.

Need to decise where to put this example. Do we wan to intoduce noisy function at this stage or we put it here because they are insights draw from the interpolating plots.

xxx

@@

*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem. Figure \ref{fig:trace-compare} shows the interpolating plots for two different searching methods: search_geodesic and serach_better on a 2D projection problem for the larger dataset using the holes index. The two methods reaches different ending value in the red dots, which is not ideal. This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.

```{r trace-compare}
p1 <- holes_2d_geo %>% 
  explore_trace_interp(col = tries) + ggtitle("search_geodesic") + ylim(0.8, 0.96)

p2 <- holes_2d_better_max_tries %>% 
  explore_trace_interp(col = tries) + ggtitle("search_better") + ylim(0.8, 0.96)

`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a larg number of basis (`n_sample`) in the neightbourhood, whose radius is controlled by `polish_alpha`. Among the `n_sample` basis, the one with the largest index value becomes the candidate. If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration. If no basis is found to have larger index value than the current basis, the searching neighbourhood will be shrunk and the search continues. The polishing search ends when one of the four stopping criteria is satisfied:

@@

```{r polish, fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish. Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}

@@

<!-- *Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in serach_better & search_posse. -->

@@

```{r alpha}
# p1 <- holes_2d_better_max_tries %>%
#   filter(info != "interpolation") %>% 
#   mutate(id = row_number()) %>% 
#   explore_trace_parameter(var = alpha) +
#   ylim(0.46, 0.505) + ggtitle("search_better")
# 
# p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
# 
# p2 <- holes_2d_pos %>%
#   filter(info != "interpolation") %>% 
#   mutate(id = row_number()) %>% 
#   explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")
# 
# p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
# 
# (p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
```

@@

Matrix parameter could also be interested to explore, for example, in the tour, we are interested to explore the relative position of the projection basis in the vector space. A projection basis is a vector/matrix whose row number is the number of variable in the given dataset. This imposes difficulties in visualisation since we are bounded to preceive at most three dimensions. Thus, principal component analysis is used to reduce the dimension of the projection basi and the first two principal component are mapped to the x and y axis of the plot. Another variable of interest could be mapped to the color aesthetics to see how it changes as the projection basis changes in the projected 2D space. [have a second read here]

@@

```{r pca, fig.cap = "\\label{pca}this is xxx"}
holes_1d_geo %>% explore_proj_pca(col = info)

@@

*Example: understand the polishing alpha via mapping alpha to color* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhod at each loop and an appropriate inital value in neighbourhood parameter alpha would avoid search on large vector space and concentrate on the ending basis. In Figure, two different initial alpha value is used in the polishing step. [some more illustration of the comparison of the two polish alpha here]

@@

```{r pca-animated, echo = TRUE, eval = FALSE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info)
```

>>>>

```{r load-data}
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

source(here::here("source/indice.r"))
source(here::here("source/sim_data.r"))
```

@@

Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-dimension projection of multivariate data. A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine. The projection pursuit index measures the "interestingness" of data defined as its departure from normality. Numerous indices have been proposed in the literature, including lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection]. An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index. The discussion of existing optimisation procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path. An illustration modified from [@buja2005computational] entails the geodesic interpolation between the plane generated by projection pursuit.

@@

As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the computational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe. An ideal optimisation procedure needs to have following characteristics:

- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentiable

- *Being able to optimise with constraints*: the projection matrix is restricted to an orthonormal matrix.

@@

@posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis. The neighbourhood is defined via the radius of the p-dimensional sphere, $c$. The new basis is taken as the target basis if it has higher index value, or the sampling continues. If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved. The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number. [Pursuit package uses this method and it works great! But I don't think we implement this - although don't think it is too hard to do it].

@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable. Since some indices could be non-differentiable, the computation of derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood. Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.

@@

*Think about how does your package help people to understand optimisation*

 - diagnostic on stochastic optim
 - vis the progression of multi-parameter decision variable 
 - understanding learning rate - neighbourhood parameter
 - understand where the local & global maximum is found - trace plot - see if noisy function

@@

The pseudo-code below illustrates the implementation of guided tour in the tourr package. Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path. This information is then used to compute a series of interpolating bases inside the `tour()` function. All the basis will be sent to create animation for visualising the tour in the `animate()` function.

@@

Visualisation has been widely used for exploring and understanding data. Visualisation presents information in a graphical manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation. Visual diagnostics can be real-time or post-run. Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.   This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.

@@

The graphical system in R is established based on the conception grammar of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.  A layer includes 1) the dataset that powers the plot; 2) a geometric object that represents the visual shape of the plot and 3) relevant statistical transformation if needed. An important concept in the grammar of graphic is *aesthetic mapping*. Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object. For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot. To create a boxplot, we first map one variable on the x-axis and then map the five point summary of another variable on the y-axis. This computation of five point summary from the origin variable is statistical transformation. This definition of graphic through layers provides advantages to produce complex plots since all the plots, however complex, can be decomposed into basic geometric objects shaped by the variable supplied.

@@

The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm. The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest. The next section shows how a global object is created in projection pursuit guided tour.

@@

In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored. This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**

Two set of simulated data are used in the demonstration of the visualisation and diagnostics of the tour optimisation. A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`). `x2` is the informative variable simulated from two bi-modal normal distribution centred at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`. The data has been scaled to ensure `x2` has variance of 1.

@@

Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure. The global structure consists of six columns: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process. The example below presents the global object of a 1D projection of the small dataset with geodesic searching method.

@@

Given all the information of interest in a global object, the next thing is to design its visual representation that can diagnose our algorithm. In an iterative algorithm, one would be interested to know how one parameter changes as the algorithm iterates. The parameter could be could be the decision variable, the value of the optimisation function or other parameter that facilitate the algorithm. In projection pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`). These are the variables that will be mapped into the diagnostic plot. The global object also has its time series nature since the observations are recorded as the algorithm progresses and thus can't be switch in rows. This time series structure is always hierarchical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm. Take our example, in the global object, each row has an `id` label which is the smallest unit the observation is ordered. A larger ordering unit is `tries` since it is updated every time a new target basis is found. A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`. These features explain the definition of the global object for projection pursuit guided tour in Figure \ref{fig:glb-obj}.

@@

@@

dt <- holes_2d_better_max_tries %>%
  filter(info != "interpolation") %>% 
  mutate(id = row_number())

@@

Plotting the interpolating points as time series data allows us to diagnose characteristics of different configurations and index functions. Here we present two examples of using  plots to diagnose the tour algorithm and different index functions.

@@

*Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index. As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to have a higher index value. After the interpolation, the target basis will become the current basis and send back to the projection pursuit algorithm to find the next target basis. From figure \ref{fig:interruption}, it is possible that there are bases with index value higher than the target basis on the interpolation path and these bases could be used to search for new basis in the next iteration.

Thus an interruption is constructed to accept the interpolating bases up to the one with the larger index value on the interpolation path, and that basis is taken as the current basis for the next iteration. After implementing this interruption, the tracing plot with the same configuration is shown on the lower panel. We can observe that rather than interpolating to the target basis at `id = 62`, the interpolation stops before the index value starts to decrease at `id = 60`. This implementation results in a higher index value in the end with fewer steps.

@@

*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{fig:interruption}, is smooth, while this may not be the case for more complicated index function. `kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF). Several visualisations below shows some characteristic of this index function.

Figure \ref{fig:geodesic-better} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better`. Rather than a smooth interpolation, the interpolation path for `kol_cdf` shows a zig-zag pattern. Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method.

```{r geodesic-better,  fig.asp=0.5, fig.cap = "Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisy index: kol_cdf. The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
p1 <- bind_rows(kol_cdf_1d_geodesic, kol_cdf_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")

p2 <- bind_rows(kol_cdf_1d_better, kol_cdf_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")

(p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
```

Moving from a 1D projection on a dataset with only one informative variable, we now introduce another informative variable in the data and remain 1D projection. In this problem, there are two local maximum being `r calc_kol_cdf(matrix(c(0, 1, 0 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 1, 0, 0, 0, 0]$ and `r calc_kol_cdf(matrix(c(0, 0, 1 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 0, 1 ,0, 0, 0]$. As in Figure \ref{fig:1d-2var-different-seeds}, using `search_better` with different seeds, both local maximum can be found. A simulated annealing algorithm, `search_better_random` in Figure \ref{fig:1d-2var-better-random}, can always find the global maximum at a cost of a large number of `tries`.

```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing). Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp() + ggtitle("search_better, seed 123456")
p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp() +  ggtitle("search_better, seed 12345")

(p5 | p6) & theme(legend.position = "none") & ylim(0, 0.25)
```

```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
kol_cdf_1d_2var_better_random %>% 
  explore_trace_interp() + 
  ylim(0, 0.25) + 
  theme(legend.position = "none") + 
  ggtitle("search_better_random")
```

@@

*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem while this may not be the case in real application. This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.

`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`. Among the `n_sample` basis, the one with the largest index value becomes the candidate. If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration. If no basis is found to have larger index value than the current basis, the searching neighbourhood will be shrunk and the search continues. The polishing search ends when one of the four stopping criteria is satisfied:

@@

```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish. Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}

@@

*Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in search_better & search_posse.

@@

```{r alpha, fig.asp=0.5,eval = FALSE}
p1 <- holes_2d_better_max_tries %>%
  filter(info != "interpolation") %>%
  mutate(id = row_number()) %>%
  explore_trace_parameter(var = alpha) +
  ylim(0.46, 0.505) + ggtitle("search_better")

p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)

p2 <- holes_2d_pos %>%
  filter(info != "interpolation") %>%
  mutate(id = row_number()) %>%
  explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")

p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)

(p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
```

@@

Matrix parameter could also be interested to explore, for example, in the tour, we are interested to explore the relative position of the projection basis in the vector space. A projection basis is a vector/matrix whose row number is the number of variable in the given dataset. This imposes difficulties in visualisation since we are bounded to perceive at most three dimensions. Thus, principal component analysis is used to reduce the dimension of the projection basis and the first two principal component are mapped to the x and y axis of the plot. Another variable of interest could be mapped to the color aesthetics to see how it changes as the projection basis changes in the projected 2D space. [have a second read here]

@@

```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic Coloring by info allows for better understanding of each stage in the geodesic search"}

holes_1d_geo %>% explore_proj_pca(col = info) + 
  theme(legend.position = "bottom")

@@

*Example: initial value for polishing alpha in search geodesic* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhood at each loop. Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value. A default initial value for polishing step is 0.5 and we are interested in whether this is an appropriate initial value. The problem is a 1D projection of the small dataset using `search_geodesic` and followed by `search_polish`. Figure \ref{fig:polish-alpha} display all projection bases on the first two principal components, colored by the `polish_alpha` parameter in the polishing step. Rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, it searches a much larger vector space that are unnecessary. Thus a user-supplied smaller initial value for `polish_alpha` would be ideal.

On the right panel of Figure \ref{fig:polish-alpha-tracing}, the `polish_alpha` is initialised as the projection distance between the last two target bases and we observe a concentrated searching space near the ending basis. With the supplied initial value for `polish_alpha`, the algorithm reaches the same ending index value with fewer iterations.

```{r polish-alpha, fig.cap = "PCA plot of two different polish alpha initialisations. A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis."}
p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
  explore_proj_pca(col = as.factor(alpha))

p2 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
  explore_proj_pca(col = as.factor(alpha))

p3 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>%
  explore_trace_interp(col = method) + ylim(0.75, 0.95)

p4 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
  explore_trace_interp(col = method) +  ylim(0.75, 0.95)

(p1 | p2) & theme(legend.position = "none")

```{r polish-alpha-tracing, fig.asp = 0.5,fig.cap="Comparing the trace of two different polish alpha initialisations. Both intialisations reach the same ending index values "}
(p3 | p4) & theme(legend.position = "none")

@@

```{r pca-animated, echo = TRUE, eval = FALSE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) + 
  theme(legend.position = "bottom")
```
